Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-20 doc/GMP2015/appendix_a.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!
Discussion of algorithm refined. Added some Meera stuff.

\section{Appendix}
\label{sec:appendix}

@@

\begin{proof}
Remark \ref{lemma:union_intersection} shows that the intersection of any two well-constrained subgraphs (such as $C_i$ and $C_j$) can only result in trivial or well-constrained subgraphs. Therefore, these are the only possibilities to consider.

@@

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is well-constrained. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a well-constrained subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be well-constrained. As $C$ is well-constrained, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.

\medskip\noindent
\vemph{Part 2:} If $C_i \cap C_j$ is well-constrained, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also well-constrained. This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself. Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

@@

For (2), observe that under-constrained subgraphs of well-constrained graphs must have density less than $k$. For (3), observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.

For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
\end{proof}

@@

%     % \item Observe that, given a well-constrained graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.

%     \item Observe that, given a well-constrained graph, a subgraph with density $k$ must also be well-constrained. Then it is straightforward application of the equation above.

%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of well-constrained $F$, it must be that $d(F_i\cup F_j)=l<k$. Therefore $d(F_i\cap F_j)=2k-l>k$. This means $F_i\cap F_j$ is trivial. \textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial. By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.

%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$. Being a subgraph of well-constrained $F$, $F_i\cap F_j$ is also well-constrained. \textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is well-constrained.

%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
% \end{enumerate}
% \end{proof}

\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_is_C}}

\begin{proof}
Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
%
In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem). Thus, $C_i\cup C_j=C$ is well-constrained.
\end{proof}

\subsubsection{Proof of Lemma \ref{lemma:wc_intersection_makes_all_wc}}

\begin{corollary}\label{corollary:no_edges_between_diff}
Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is well-constrained, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
\end{corollary}

@@

\begin{proof} (Of Lemma \ref{lemma:wc_intersection_makes_all_wc})

@@

Then we assume that there is some third well-constrained vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

@@

    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Corollary \ref{corollary:no_edges_between_diff}) and cannot be well-constrained.

    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be well-constrained. We also know it cannot be trivial because it contains well-constrained subgraphs. This means it must be under-constrained. From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial. This is impossible because $D_{i,j}$ is well-constrained, thereby contradicting the assumption that $C_k$ is well-constrained.

@@

\subsubsection{Proof of Lemma \ref{lemma:uc_intersection_makes_all_uc}}

\begin{proof}
Assume there is some $k$ such that $C_i\cap C_k$ is not trivial. By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be well-constrained. Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be well-constrained. This means that $C_i\cap C_j$ is well-constrained. Therefore, such a $k$ cannot exist and all intersections are trivial.
\end{proof}

@@

First notice that since $C_i \cup C_j$ is well-constrained, by Observation~\ref{lem:pinned_union_intersection},
both $C_i$ and $C_j$ are connected well-constrained vertex-maximal proper subgraphs of  $C$.
Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a well-constrained subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected well-constrained vertex-maximal proper subgraph of  $C$.

Then we analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be well-constrained.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are well-constrained.

>>>>

\section{Appendix A}
\label{sec:appendix:a}

@@

\begin{proof}
Remark \ref{lemma:union_intersection} shows that the intersection of any two isostatic subgraphs (such as $C_i$ and $C_j$) can only result in trivial or isostatic subgraphs. Therefore, these are the only possibilities to consider.

@@

Take the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ which we \vemph{assume} is isostatic. If $U\neq C$, then we just found a larger proper subgraph and the children were not vertex-maximal to begin with. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial subgraphs. By definition of a DR-plan $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is both one or more trivial subgraphs and a isostatic subgraph of $C$. Thus, the assumption is wrong and $U$ cannot be isostatic. As $C$ is isostatic, this means no proper subset of the children in ComDRP$(C)$ can union to form $C$.

\medskip\noindent
\vemph{Part 2:} If $C_i \cap C_j$ is isostatic, then, by Remark \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, any two children of $C$ will union to $C$ itself. Thus, any two children are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

@@

For (2), observe that under-constrained subgraphs of isostatic graphs must have density less than $k$. For (3), observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic. Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.

For (4), because subgraphs of a isostatic graph can only be trivial, under-, or isostatic, all cases have already been exhausted.
\end{proof}

@@

%     % \item Observe that, given a isostatic graph, an under-constrained subgraph must have density less than $k$. Then it is straightforward application of the equation above.

%     \item Observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic. Then it is straightforward application of the equation above.

%     % \item \textit{Forward direction:} Since $F_i\cup F_j$ is under-constrained and a subgraph of isostatic $F$, it must be that $d(F_i\cup F_j)=l<k$. Therefore $d(F_i\cap F_j)=2k-l>k$. This means $F_i\cap F_j$ is trivial. \textit{Reverse direction:} We know that $d(F_i\cap F_j)>k$ because it is trivial. By the same math, we find that $d(F_i\cup F_j)<k$, showing it is under-constrained.

%     % \item \textit{Forward direction:} We have that $d(F_i\cup F_j)=k$, therefore $d(F_i\cap F_j)=k$. Being a subgraph of isostatic $F$, $F_i\cap F_j$ is also isostatic. \textit{Reverse direction:} By the same math, we find that $d(F_i\cup F_j)=k$ and, since it is also a subgraph of $F$, it is isostatic.

%     \item Subgraphs of a isostatic graph can only be trivial, under-, or isostatic. All cases are already exhausted.
% \end{enumerate}
% \end{proof}

\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}}

\begin{proof}
Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
%
In the reverse direction, we know $C$ is either some non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
\end{proof}

\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}}

\begin{observation}\label{observation:no_edges_between_diff}
Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
\end{observation}

@@

\begin{proof} (Of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc})

@@

Then we assume that there is some third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

@@

    \item 4 cases: \usestwod $C_k$ cannot be $\inducedOnC{R_i\cup R_j}$, $\inducedOnC{R'_i\cup R_j}$, $\inducedOnC{R_i\cup R'_j}$, or $\inducedOnC{R'_i\cup R'_j}$ because these are all disconnected (Observation \ref{observation:no_edges_between_diff}) and cannot be isostatic.

    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_i\cup C_k = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_i\cup C_k$ cannot be isostatic. We also know it cannot be trivial because it contains isostatic subgraphs. This means it must be under-constrained. From Lemma \ref{lemma:union_intersection}, we know that $C_i\cap C_k=\inducedOnC{R'_j\cup D_{i,j}}$ must then be trivial. This is impossible because $D_{i,j}$ is isostatic, thereby contradicting the assumption that $C_k$ is isostatic.

@@

\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:uc_intersection_makes_all_uc}}

\begin{proof}
Assume there is some $k$ such that $C_i\cap C_k$ is not trivial. By Remark \ref{lemma:union_intersection}, $C_i\cap C_k$ must be isostatic. Then, by Lemma \ref{lemma:wc_intersection_makes_all_wc}, the intersection between any two children must be isostatic. This means that $C_i\cap C_j$ is isostatic. Therefore, such a $k$ cannot exist and all intersections are trivial.
\end{proof}

\subsubsection{Proof of Theorem \ref{theorem:algo_complexity}}

\begin{proof}
The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be the list of component-candidates. Then run the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} \todo{complexity?} on each candidate with edge $e$, which will find the minimal subgraph containing the candidate and $e$, which we call $D$. If $D=G$, move $D$ to the true-component list and remove the candidate from the candidate-component list. If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list. Then, check the union of $D$ with the all items in the true-component list; if the union is isostatic, halt and output just these two components as children. Otherwise, run to completion and output all items in the true-component list.

This is run recursively on each node. If you use dynamic programming to store the decompositions of each component, it prevents repetition of work. In this case you will have to do work on no more than $O(V)$ nodes that get smaller at each level.

Thus, the running time is at most $O(|V|^3)$. \todo{does this amortize correctly?}
\end{proof}

% \begin{proof}
% The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones. After the subgraphs have been found, it is a straightforward application of the theory. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs. The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time). \todo{Think more about complexity.}
% \end{proof}

@@

First notice that since $C_i \cup C_j$ is isostatic, by Observation~\ref{lem:pinned_union_intersection},
both $C_i$ and $C_j$ are connected isostatic vertex-maximal proper subgraphs of  $C$.
Since $C_j \cup C_j = C$, there are no edges in $C$ that is not contained in a isostatic subgraph, so $C$ does not have any single-edge child node, and $C_k$ is a connected isostatic vertex-maximal proper subgraph of  $C$.

Then we analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=Idc(C,R'_i\cup R'_j\cup D_{i,j})$ is not possible. Since $C_k\cup C_i = Idc(C,R'_i\cup R_j\cup D_{i,j})\neq C$ we have from Lemma \ref{lemma:wc_intersection_is_C} that $C_k\cup C_i$ cannot be isostatic.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be under-constrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

### 2014-12-20 doc/GMP2015/appendix_b.tex
Discussion of algorithm refined. Added some Meera stuff.

>>>>

\subsection{Open Problems}

\todo{added these in newest conclusion}

A few natural  open questions concern the following common theme
that runs through the paper:

\begin{openproblem}
For fixed $k$ we have polynomial time optimal DR-planning,
recombination (modification) in the presence of $k$ overconstraints (Section
\ref{sec:DRP},
optimal modification for decomposition {\sl OMD}$_k(G)$,
when at most $k$ constraints are removed
\ref{sec:recomb}
and
also optimal completion
using at most $k\le 2$ constraints in the body-pin and
triangle-multipin cases
under a modified measure for optimizing the
DR-plan \ref{sec:table}.
However, $k$ appears in the exponent for
the straightforward algorithms for these problems.
\end{openproblem}

One problem in the above theme is from
Section \ref{sec:bodypin}.
\begin{openproblem}
    What is the complexity of the optimal completion problem when the given
    graph has more than 2-dofs?
    Our proof for the 1 and 2-dof cases relied heavily on the matroidal
    properties of their corresponding $(k,l)$-tightness.
    For higher number of dofs, the $(k,l)$ characterization is
    no longer matroidal \cite{Lee:2007:PGA}.
    As a result, the major obstacle is that
    there is no easy way of obtaining an optimal or canonical
    $k$-dof DR-plan in general.
    Even assuming such  a DR-plan is available,
    if higher dofs had the same characteristics,
    Observation \ref{obs:algebraic_completion}
    raises questions about the correct measure of DR-plan size that
    captures algebraic complexity for recombining graphs with many dofs
    (this is not an issue in the isostatic case). Unless some restrictions
    can be found and taken advantage of, the $k$-dof optimal
    completion problem would  have complexity exponential in $k$.
\end{openproblem}

Another problem from the above theme is from Section \ref{sec:recomb}
\begin{openproblem}
What is the complexity of the restricted OMD (optimal modification for
decomposition) problem?
This has the potential to be difficult. For example, when the
isostatic completion is required to be a 2-tree the restricted OMD
problem is reducible to the maximum spanning series-parallel subgraph
problem shown by \cite{cai1993spanning} to be NP-complete even if the input
graph is planar of maximum degree at most 6. However, since the OMD
problem has other input restrictions such as not having any proper
isostatic subgraphs, it is not clear if the reverse reduction exists
and hence it is unclear whether the OMD problem is NP-complete.

More general problem directions are the following.
\begin{openproblem}
    Combinatorial rigidity for
    periodic structures is an active area of research. This paper motivates
    a study of the rigidity of self-similar structures,
    with self-similar groups replacing periodic groups.
\end{openproblem}

### 2014-12-20 doc/GMP2015/s00_abstract.tex
White space fixes.
Discussion of algorithm refined. Added some Meera stuff.

\begin{abstract}
A variety of quasi-uniform (aperiodic) and possibly self-similar,
layered material structures are a natural consequence of design
objectives such as rigidity or flexibility, minimizing mass, optimally
distributing external stresses and participating in the assembly of
diverse and multifunctional, larger  structures.

Crucial for the design and analysis of the abovementioned properties
is the recursive decomposition of diverse types of underlying 2D
geometric constraint systems that are \vemph{isostatic or
underconstrained (which we call collectively as \dfn{qusecs})}. While
the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
hard even for (possibly overconstrained) 2D bar-joint constraint
systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
plans that meet various desirable, previously studied criteria. We
leverage recent results on Cayley configuration spaces to show that
the indecomposable systems -  that are solved at node of the optimal
DR-plan by recombining solutions to child systems - can be minimally
modified to become decomposable and have a small DR-plan, leading to
efficient realization algorithms. We show formal connections to well-
known problems such as completion of underconstrained systems. We
briefly discuss the modeling and analysis of two specific 2D layered
materials as applications of the above theory and algorithms. We
intend to make a software implementation and videos available upon
request and publicly available for the final version.

@@

% % \item
% % Software implementation and videos can be found at: \todo{Provide Link}.
\end{abstract}

>>>>

\begin{abstract}
A variety of quasi-uniform (aperiodic) and self-similar,
layered material structures are a natural consequence of design
objectives such as rigidity or flexibility, minimizing mass, optimally
distributing external stresses and participating in the assembly of
diverse and multifunctional, larger  structures.

Crucial for the design and analysis of the abovementioned properties
is the recursive decomposition of diverse types of underlying 2D
geometric constraint systems that are \vemph{isostatic or
underconstrained (which we call collectively as \dfn{qusecs})}. While
the optimal recursive decomposition (\dfn{DR-planning}) problem is NP-
hard even for (possibly overconstrained) 2D bar-joint constraint
systems we describe an $O(n^2)$ algorithm  for qusecs. Besides
elucidating NP-hardness, the algorithm outputs \dfn{canonical} DR-
plans that meet various desirable, previously studied criteria. We
leverage recent results on Cayley configuration spaces to show that
the indecomposable systems -  that are solved at node of the optimal
DR-plan by recombining solutions to child systems - can be minimally
modified to become decomposable and have a small DR-plan, leading to
efficient realization algorithms. We show formal connections to well-
known problems such as completion of underconstrained systems. We
briefly discuss the modeling and analysis of two specific 2D layered
materials as applications of the above theory and algorithms. We
intend to make a software implementation and videos available upon
request and publicly available for the final version.
\end{abstract}

@@

% % \item
% % Software implementation and videos can be found at: \todo{Provide Link}.

### 2014-12-20 doc/GMP2015/s01_introduction.tex
I did it! Lots of little changes and tweeks to white space. It's in 10 pages!
Discussion of algorithm refined. Added some Meera stuff.

Examples of such materials (See Figure \ref{fig:material_examples})
include:

\begin{enumerate}
    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}

    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}) \uncited

@@

  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems.

@@

>>>>

Examples of such materials (See Figure \ref{fig:material_examples})
include:
%
\begin{enumerate}
    \item Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.

    \item Cross-sections of organic tissue with hierarchical structure, e.g. compact bone and tendon (Figure \ref{fig:material_examples:tendon}).

@@

  \item We give an efficient ($O(n^2)$) algorithm to find a canonical
  (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned
  above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and
  \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence
  of the NP-hardness of finding optimal DR-plans for over-constrained
  systems. \todo{wording}
  Our optimal/canonical DR-plan satisfies desirable properties such as cluster minimality previously studied \cite{hoffman2001decompositionI}.

@@

\note
All proofs are in Appendix A (\ref{sec:appendix:a}).

### 2014-12-20 doc/GMP2015/s03_drp.tex
Discussion of algorithm refined. Added some Meera stuff.

\begin{lemma}
Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
Then
\begin{enumerate}
    \item
    \label{lemma:wc_intersection_is_C}
$C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.
\item
    \label{lemma:wc_intersection_makes_all_wc}
If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic
Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.
\item
    \label{lemma:uc_intersection_makes_all_uc}
If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{enumerate}
\end{lemma}

@@

\subsection{Algorithm}

The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \cite{Oung:2001:FFE:376957.376995} to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \cite{Jacobs:1997:PG} on each node and taking the largest components that are not subgraphs of any larger ones. After the subgraphs have been found, it is a straightforward application of the theory. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs. The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time). \todo{Think more about complexity.}

>>>>

\begin{lemma}\label{lemma:combined_lemma}
Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
Then
\begin{enumerate}
    \item\label{lemma:wc_intersection_is_C}
    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

    \item\label{lemma:wc_intersection_makes_all_wc}
    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.

    \item\label{lemma:uc_intersection_makes_all_uc}
    If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{enumerate}
\end{lemma}

@@

% \subsection{Algorithm}

\begin{theorem}\label{theorem:algo_complexity}
    There exists an $O(|V|^3)$ algorithm to find a canonical DR-plan for any 2D isostatic bar-joint graph.
\end{theorem}

\begin{remark}
    while the canonical DR-plan is optimal only if the input graph is independent, if it has only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan. However, the time complexity is exponential in $k$.
\end{remark}

### 2014-12-20 doc/GMP2015/s09_conclusion.tex
Discussion of algorithm refined. Added some Meera stuff.

\section{Conclusion}
\label{sec:conclusion}
% \bibliography{paper}

We have elucidated the complexity barriers, for example over-constraints, faced by the general optimal DR-plan and recombination problem. When there are no overconstraints, which is the case for 2D qusecs whose realizations are many common types of layered materials, we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property given an efficient ($O(n^2)$) algorithm to find a DR-plan that satisfies other desirable characteristics.

We have also described novel methods of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise non-decomposable system that recombines the solved child systems at a node of a DR-plan. These results rely on a recent theory of Convex Cayley Configuration spaces.

>>>>

\section{Conclusion}
% \bibliography{paper}

We have clarified the main source of complexity for the optimal DR-plan and recombination problems. For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient ($O(n^2)$) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.

We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan These results rely on a recent theory of Convex Cayley Configuration spaces. Relationships and reductions between these and previously studied problems were formally clarified.
